---
title: "IMDB_Analysis_Group_1_Appendix"
output: html_document
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
library(ggplot2)
library(ggthemes)
library(ggrepel)
library(reshape2)
library(dplyr)
library(plyr)
library(caret)
library(scales)
library(countrycode)
library(GGally)
library(knitr)
library(scales)
library(gridExtra)
library(pander)
library(stargazer)
```
#Appendix


```{r, echo = FALSE}
movies <- read.csv(file = "movie_metadata.csv", header = TRUE, stringsAsFactors = FALSE, strip.white = TRUE)
```



```{r, echo = FALSE}
## To remove rows where NAs are present for any of the applicable columns
movies <- movies[complete.cases(movies[c("title_year","budget","gross")]),]
## To remove aspect ratio and imdb link columns
movies$aspect_ratio <- NULL
movies$movie_imdb_link <- NULL
```


```{r, echo = FALSE}
## for loop to replace NAs with means for particular columns
for (i in c("actor_1_facebook_likes" , "actor_2_facebook_likes" , "actor_3_facebook_likes" , "num_critic_for_reviews" , "duration" , "facenumber_in_poster")){
    k <- which(colnames(movies) == i)
    movies[k][is.na(movies[k] == TRUE)] <- round(mean(movies[[k]], na.rm = TRUE), 0)
}
```

```{r, echo = FALSE}
movies <- movies[!duplicated(movies$movie_title),]
# Function to remove Â, leading and trailing whitespace from movies$movie_title
movie_title_processing <- function(str){
  str <- sub(pattern = "Â", replacement = "", str)
  str <- sub(pattern = "^\\s+|\\s+$", replacement ="", str)
}
# Apply previous function
movies$movie_title <- sapply(movies$movie_title, FUN = movie_title_processing)
```

##Nikhita

```{r echo=FALSE}
counted <- count(movies, vars = "country")
countries <- subset(movies, country %in% c("Australia", "Canada", "China", "France", "Germany", "Hong Kong", "India", "Italy", "Japan", "Spain", "UK", "USA"))
score <- countries$imdb_score
topratedmovies <- subset(movies, content_rating %in% c("PG-13", "R", "PG"))
contentrating <- topratedmovies$content_rating
contentscore <- topratedmovies$imdb_score
contentgross <- topratedmovies$gross
ggplot(data = counted, aes(x = country, y = freq, fill = country)) + 
    geom_bar(stat = "identity", colour = "black") +
    labs(title = "Number of movies by country", x = "", y = "") + 
    geom_text(aes(label = freq), hjust = -0.2, vjust = 0.4) + 
    coord_flip() +
    theme_few() +
    theme(legend.position = "None") +
    theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.ticks.y = element_blank(),
          axis.title.x = element_blank(), axis.title.y = element_blank()) 

# Plot of imdb score for the top 12 countries
ggplot(countries, aes(country, score, fill = country)) + 
    geom_boxplot() + stat_summary(fun.y=mean, colour="red", geom="point", size=1, show.legend = FALSE) + labs(title = "IMDB Score of movies by country", x = "", y = "Gross revenue") + 
    labs(title = "IMDB Score of movies by country", x = "", y = "IMDB Score") + 
    theme_bw() +
    theme(legend.position = "None") +
    theme(axis.ticks.y = element_blank(), panel.grid.major.y = element_blank(), axis.title.x = element_blank()) +
    scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10), limits = c(1, 10))

# Looking at the IMDB scores, movies across all ten countries have a mean score between 6 and 7. In India & Japan, at least 40% of the movies scored more than 7, whereas movies in Australia and USA have almost equal numbers scoring between 6 and 7.

## IMDB Score analysis for top three ratings
# Boxplot illustration
ggplot(topratedmovies, aes(contentrating, contentscore, fill = contentrating)) + 
    geom_boxplot() + stat_summary(fun.y=mean, colour="red", geom="point", size=1, show.legend = FALSE) +
    labs(title = "IMDB Score of movies by rating", x = "Content Rating", y = "IMDB Score") + 
    theme_bw() +
    scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10), limits = c(1, 10)) +
    theme(legend.position = "none", axis.ticks.x = element_blank(), panel.grid.major.x = element_blank())

# Does the mean IMDB Score vary with different ratings?
imdbscore <- movies$imdb_score
contentrating <- movies$content_rating

scoret1 <- t.test(imdbscore[contentrating == "R"], imdbscore[contentrating == "PG-13"])
pander(scoret1, caption="")

scoret2 <- t.test(imdbscore[contentrating == "PG-13"], imdbscore[contentrating == "PG"])
pander(scoret2, caption="")

scoret3 <- t.test(imdbscore[contentrating == "R"], imdbscore[contentrating == "PG"])
pander(scoret3, caption="")
```

##Steven

Appendix
```{r, echo=FALSE, fig.width = 10, fig.align = "center", message = FALSE}

# imdb score vs year
temp <- movies[!is.na(movies$title_year),]
ggplot(temp, aes(x = factor(title_year), y = imdb_score, fill = factor(title_year))) + 
  geom_boxplot(outlier.shape = NA) + stat_summary(fun.y=mean, colour="red", geom="point", size=1, show.legend = FALSE) +
  theme(axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5),legend.position = "None") +
  labs(title = "IMDB scores against year", x = "", y = "IMDB Score") +
  scale_y_continuous(limits = c(2.5, max(movies$imdb_score)))
```

The plot above shows that there might be a downward trend in IMDB scores over the years. Some years, however, include fewer than ten movies. To get a more accurate picture of the development of IMDB scores only years with ten or more entries will be plotted below.

```{r, echo=FALSE, fig.width = 10, fig.align = "center", message = FALSE}
# subset movies data frame to include all movies from 1980 onwards
temp <- movies[movies$title_year >= 1980,]
# plot the range of IMDB scores for all years
p1 <- ggplot(temp, aes(x = factor(title_year), y = imdb_score, fill = factor(title_year))) + 
  geom_boxplot(outlier.shape = NA) + stat_summary(fun.y=mean, colour="red", geom="point", size=1, show.legend = FALSE) +
  theme(axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5),legend.position = "None") +
  labs(title = "IMDB scores against year, for years with >= 10 entries", x = "", y = "IMDB Score") +
    geom_boxplot(outlier.shape = NA) +
    theme_bw() +
    theme(axis.ticks.y = element_blank(), panel.grid.major.y = element_blank(), axis.title.x = element_blank(),
          axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5),legend.position = "None") +
    labs(title = "IMDB scores against year, for years with >= 10 entries", x = "", y = "IMDB Score") +
    scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10), limits = c(1, 10))

lm_model <- lm(imdb_score ~ title_year, data = temp)

# make a scatter plot that shows trend in IMDB scores
p2 <- ggplot(temp, aes(x = title_year, y = imdb_score, colour = factor(title_year))) + 
    geom_jitter(alpha = 0.4) +
    geom_abline(intercept = lm_model$coefficients[1], slope = lm_model$coefficients[2], size = 1) +
    theme_bw() +
    labs(x = "", y = "IMDB Score") +
    theme(axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5), legend.position = "None", axis.title.x = element_blank()) +
    scale_x_continuous(breaks = seq(min(temp$title_year), max(temp$title_year), by = 1), 0.5) +
    scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10), limits = c(1, 10)) 

grid.arrange(p1, p2, ncol=1)
```

The plots above give a less biased represenation of the development of IMDB scores. The scores now appear more constant over the years. Although minimal, the scatter plot reveals that there is still a slight downward trend in scores.

```{r, echo = FALSE, fig.width = 10, fig.align = "center"}
# create function that converts years into decades
convert_decade <- function(year){
  low <- year - year %% 10
  high <- year - year %% 10 + 9
  paste(as.character(low), as.character(high), sep = "-")
}
# apply previous function
movies$decade <- sapply(movies$title_year, FUN = convert_decade)
```

```{r, echo=FALSE, fig.width = 10, fig.align = "center", message = FALSE}
# for all countries by decade

# remove entries with NA-NA
temp <- movies[!movies$decade == "NA-NA", ]
ggplot(temp, aes(x = decade, y = imdb_score, fill = decade)) + 
  geom_boxplot(outlier.shape = NA) + stat_summary(fun.y=mean, colour="red", geom="point", size=1, show.legend = FALSE) +
  theme(axis.text.x=element_text(angle=45, hjust=0.5, vjust=0.5), legend.position = "None") +
  labs(title = "IMDB scores against decade, all countries", x = "", y = "IMDB Score")
```


```{r, echo=FALSE, fig.width = 10, fig.align = "center", message = FALSE}
# facetgrid of imdb scores over decade for continents
movies$continent <- countrycode(as.character(movies$country), "country.name", "continent")
south_America = c("Brazil", "Argentina", "Chile", "Colombia", "Peru")

for (i in 1:nrow(movies)) {
  if (is.na(movies$continent[i])) {
    next
  } else if ((movies$continent[i] == "Americas") & (movies$country[i] %in% south_America)) {
    movies$continent[i] <- "South America"
  } else if ((movies$continent[i] == "Americas") & (!(movies$country[i] %in% south_America))) {
    movies$continent[i] <- "North America"
  } 
}
temp <- movies[!is.na(movies$continent), ]
ggplot(temp, aes(y = imdb_score, x = factor(decade), fill = decade)) + 
    geom_boxplot() + 
    labs(title = "IMDB and decades for continents", y = "IMDB score", x = "Decade") + 
    facet_wrap(~ continent) +
    theme_bw() + 
    theme(legend.position = "None",axis.text.x=element_text(angle=45, hjust=0.5, vjust=0.5)) +
    annotation_logticks(sides = "lr", colour = "gray")
```

```{r, echo=FALSE, fig.width = 10, fig.align = "center", message = FALSE}
# facetgrid of imdb scores over decade for Europe
temp <- movies[movies$continent == "Europe", ]
temp <- temp[!is.na(temp$country), ]
ggplot(temp, aes(y = imdb_score, x = factor(decade), fill = decade)) + 
    geom_boxplot() + 
    labs(title = "IMDB and decades for European Countries", y = "IMDB score", x = "Decade") + 
    facet_wrap(~ country) +
    theme_bw() + 
    theme(legend.position = "None",axis.text.x=element_text(angle=45, hjust=0.5, vjust=0.5)) +
    annotation_logticks(sides = "lr", colour = "gray")

```

```{r, echo=FALSE, fig.width = 10, fig.align = "center", message = FALSE}
# facetgrid of imdb scores over decade for North America
temp <- movies[movies$continent == "North America", ]
temp <- temp[!is.na(temp$country), ]
ggplot(temp, aes(y = imdb_score, x = factor(decade), fill = decade)) + 
    geom_boxplot() + 
    labs(title = "IMDB and Decades for North American Countries", y = "IMDB Score", x = "Decade") + 
    facet_wrap(~ country) +
    theme_bw() + 
    theme(legend.position = "None",axis.text.x=element_text(angle=45, hjust=0.5, vjust=0.5)) +
    annotation_logticks(sides = "lr", colour = "gray")

```


```{r, echo=FALSE, fig.width = 10, fig.align = "center", message = FALSE}
# facetgrid of imdb scores over decade for Asia
temp <- movies[movies$continent == "Asia", ]
temp <- temp[!is.na(temp$country), ]
ggplot(temp, aes(y = imdb_score, x = factor(decade), fill = decade)) + 
    geom_boxplot() + 
    labs(title = "IMDB and Decades for Asian Countries", y = "IMDB Score", x = "Decade") + 
    facet_wrap(~ country) +
    theme_bw() + 
    theme(legend.position = "None",axis.text.x=element_text(angle=45, hjust=0.5, vjust=0.5)) +
    annotation_logticks(sides = "lr", colour = "gray")

```


##Cecilia


```{r echo = FALSE, fig.width = 10, fig.align = "center"}
#scatter plot for budget vs imdb score
ggplot(data = movies, aes(y = budget, x = imdb_score)) + 
    geom_jitter(alpha = 0.4, colour = "#FF9999") + 
    theme_bw() +
    labs(y = "Budget", x = "IMDB Score", title = "Budget versus IMDB Score") + 
    scale_y_log10(breaks = c(1, 1e+02, 1e+04, 1e+06, 1e+08, 1e+10), 
                  labels = c("1", "100", "10,000", "1 million", "100 millions", "10 billions"), limits = c(1, 1e+10))+
    scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10), limits = c(1, 10)) +
    annotation_logticks(sides = "lr", colour = "gray") 

#covariance test
cov(movies$budget, movies$imdb_score)

#correlation test
pander(cor.test(movies$budget, movies$imdb_score), caption="")
```

The covariance value of 5.3x10^7 suggests that there is an upward trend - as budget increases, imdb score increases. The Pearson's correlation test between budget and IMDB score gives a p-value of 0.047, which suggests that there isn't a high correlation between budget of the movie and rating on IMDB. 

Even if we dig deeper to look at the effect of budget on IMDB score for movies from different countries, it appears to be that there is not much of a trend.

```{r echo = FALSE, fig.width = 10, fig.align = "center"}
ggplot(budgetcountry, aes(y = budget, x = imdb_score, colour = country)) + 
    geom_jitter(alpha = 0.2) + 
    labs(title = "Budget and IMDB score for different countries", y = "Budget", x = "IMDB Score") + 
    scale_y_log10(breaks = c(1, 1e+02, 1e+04, 1e+06, 1e+08, 1e+10), 
                  labels = c("1", "100", "10,000", "1 million", "100 millions", "10 billions"), limits = c(1, 1e+10)) +
    scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10), limits = c(1, 10)) +
    facet_wrap(~ country) +
    theme_bw() + 
    theme(legend.position = "None") +
    annotation_logticks(sides = "lr", colour = "gray") 
```


```{r echo=FALSE}

#Analysis on the effect of budget on gross profit margin

#create grossprofit column
movieswprofit <- data.frame(movies)
movieswprofit$grossprofitmargin <- (movieswprofit$gross - movieswprofit$budget)/movieswprofit$gross * 100

#scatter plot for budget vs gross profit margin 
ggplot(data = movieswprofit) + geom_point(aes(x = budget, y = grossprofitmargin, colour = "#FF9999")) + scale_x_log10(breaks = c(1e+03, 1e+05, 1e+07, 1e+09), labels = c("1000", "100,000", "10 million", "1000 million")) + theme(legend.position="none") + scale_y_continuous(limits = c(-1000, 100)) + theme(legend.position="none") + labs(x = "Budget in USD", y = "Gross Profit Margin as %")

#covariance test

cov(movieswprofit$budget, movieswprofit$grossprofitmargin, use = "complete.obs") 

#correlation test

cor.test(movieswprofit$budget, movieswprofit$grossprofitmargin, use = "complete.obs")

movieswprofit$profitable <- ifelse(movieswprofit$grossprofitmargin >= 0, "yes", "no")
table(movieswprofit$profitable)

```

```{r, echo = FALSE}
subsetcountry <- count(movies, vars = "country") 
subsetcountry <- subsetcountry[order(subsetcountry[,2], decreasing = FALSE),] 
subsetcountry$country <- factor(subsetcountry$country, levels = subsetcountry$country)
subsetcountry <- subsetcountry[subsetcountry[,2] >= 10,]

#plot for budget vs imdb score for countries
budgetcountry <- subset(movies, country %in% c("Australia", "Canada", "China", "France", "Germany", "Hong Kong", "India", "Italy", "Japan", "Mexico", "Spain", "UK", "USA"))
budgetcountry <- subset(movies, country %in% subsetcountry$country)
budgetcountry$country <- factor(budgetcountry$country, levels = subsetcountry$country, ordered = TRUE)
ggplot(data = budgetcountry) + geom_jitter(aes(x = budget, y = imdb_score, color=country, alpha = 0.1)) + labs(x = "Budget", y = "IMDB Score") + scale_x_log10(breaks = c(1e+03, 1e+05, 1e+07, 1e+09), labels = c("1000", "100,000", "10 million", "1000 million"))
```

```{r, echo = FALSE}

#plot for budget vs gross for countries

ggplot(data = budgetcountry) + geom_jitter(aes(x = budget, y = gross, color=country, alpha = 0.1)) + labs(x = "Budget", y = "Gross Revenue") + scale_x_log10() 

```


We would like to investigate whether the imdb score of movie is affected by the budget for the particular movie.

```{r echo = FALSE, fig.width = 10, fig.align = "center"}
#scatter plot for budget vs imdb score
ggplot(data = movies, aes(y = budget, x = imdb_score)) + 
    geom_jitter(alpha = 0.4, colour = "#FF9999") + 
    theme_bw() +
    labs(y = "Budget", x = "IMDB Score", title = "Budget versus IMDB Score") + 
    scale_y_log10(breaks = c(1, 1e+02, 1e+04, 1e+06, 1e+08, 1e+10), 
                  labels = c("1", "100", "10,000", "1 million", "100 millions", "10 billions"), limits = c(1, 1e+10))+
    scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10), limits = c(1, 10)) +
    annotation_logticks(sides = "lr", colour = "gray") 

#covariance test
cov(movies$budget, movies$imdb_score)

#correlation test
cor.test(movies$budget, movies$imdb_score)
```

The covariance value of 5.3x10^7 suggests that there is an upward trend - as budget increases, imdb score increases. The Pearson's correlation test between budget and IMDB score gives a p-value of 0.047, which suggests that there isn't a high correlation between budget of the movie and rating on IMDB. 

Even if we dig deeper to look at the effect of budget on IMDB score for movies from different countries, it appears to be that there is not much of a trend.

```{r echo = FALSE, fig.width = 10, fig.align = "center"}
ggplot(budgetcountry, aes(y = budget, x = imdb_score, colour = country)) + 
    geom_jitter(alpha = 0.2) + 
    labs(title = "Budget and IMDB score for different countries", y = "Budget", x = "IMDB Score") + 
    scale_y_log10(breaks = c(1, 1e+02, 1e+04, 1e+06, 1e+08, 1e+10), 
                  labels = c("1", "100", "10,000", "1 million", "100 millions", "10 billions"), limits = c(1, 1e+10)) +
    scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10), limits = c(1, 10)) +
    facet_wrap(~ country) +
    theme_bw() + 
    theme(legend.position = "None") +
    annotation_logticks(sides = "lr", colour = "gray") 
```

```{r echo=FALSE}

ggplot(budgetcountry, aes(x = budget, y = imdb_score, colour = country)) + 
  geom_jitter() + scale_x_log10()+ labs(title = "Budget and IMDB score for different countries", x = "Budget", y = "IMDB Score") + scale_x_log10(breaks = c(1e+03, 1e+05, 1e+07, 1e+09), labels = c("1000", "100,000", "10 million", "1000 million")) +
  facet_wrap(~ country) +
  theme_bw() + 
  theme(legend.position = "None")

```


Does budget affect gross profit margin?

```{r echo = FALSE, fig.width = 10, fig.align = "center"}
#create grossprofit column
movieswprofit <- data.frame(movies)
movieswprofit$grossprofitmargin <- (movieswprofit$gross - movieswprofit$budget)/movieswprofit$gross * 100

#scatter plot for budget vs gross profit margin 
ggplot(data = movieswprofit, aes(x = budget, y = grossprofitmargin)) + 
    geom_point(alpha = 0.4, colour = "#FF9999") + 
    theme_bw() +
    scale_x_log10(breaks = c(1, 1e+02, 1e+04, 1e+06, 1e+08, 1e+10), 
                  labels = c("1", "100", "10,000", "1 million", "100 millions", "10 billions"), limits = c(1, 1e+10)) + 
    scale_y_continuous(limits = c(-1000, 100)) + 
    labs(x = "Budget in USD", y = "Gross Profit Margin as %") +
    annotation_logticks(sides = "tb", colour = "gray") 

#covariance test
cov(movieswprofit$budget, movieswprofit$grossprofitmargin, use = "complete.obs") 

#correlation test
cor.test(movieswprofit$budget, movieswprofit$grossprofitmargin, use = "complete.obs")
```

Finally, we would also like to look at the effect of gross vs on gross profit margin, whether the an increase in budget would make a better movie, and hence a higher gross profit margin. Gross profit margin can be used as an indictor of profitability of movie. Based on the data, it appears that there is a negative correlation (-0.31) between budget and profit. Although it is not very strong, it still suggests that an increase in budget has a slight negative impact on gross profit margin.

From the plot, we can see that a lot of movies are not profitable. Around half of the movies are profitable (count = 1988), half are not (count = 1801).

```{r echo = FALSE, fig.width = 10, fig.align = "center"}
movieswprofit$profitable <- ifelse(movieswprofit$grossprofitmargin >= 0, "yes", "no")

table(movieswprofit$profitable)
```

##Siow Meng

##George

```{r}
##Code to make this run

genres <- c()
i <- 1
for (ins in movies$genres){
    g <- strsplit(ins, "[|]")
    for (gnr in g[[1]]){
        if (!(gnr %in% genres)){
            genres[i] <- gnr
            i = i + 1
        }
    }
}
# Create a dataframe with logical values which 
# indiacte the categories of each movie
movies$genres <- strsplit(movies$genres, "[|]")
genres_idx <- movies[, c("movie_title", "genres")]
i = 1
mat <- matrix(rep(0, (dim(movies)[1] * length(genres))), nrow = dim(movies)[1])
for (g in genres_idx$genres){
    idx <- which(genres %in% g)
    mat[i, idx] <- 1
    i = i + 1
}
colnames(mat) <- genres
movies_and_genres <- data.frame(mat)

# Find how many movies belong in each genre
sum <- rep(0, length(genres))
for (i in 1:length(genres)){
    sum[i] <- sum(movies_and_genres[, i])
}
genres_sum <- data.frame(genre = factor(genres), sum = sum)
genres_sum <- genres_sum[order(sum, decreasing = FALSE),]
genres_sum$genre <- factor(genres_sum$genre, levels = genres_sum$genre)
genres_sum <- genres_sum[genres_sum$sum > 10, ]

# Number of movies belonging to each genre
ggplot(genres_sum, aes(x = genre, y = sum, fill = genre)) + 
    geom_bar(stat = "identity", colour = "black") + 
    coord_flip() +
    labs(title = "Number of movies by genre", x = "", y = "") + 
    geom_text(aes(label = sum), hjust = -0.2, vjust = 0.4) + 
    theme_few() +
    theme(legend.position = "None") +
    theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.ticks.y = element_blank(),
          axis.title.x = element_blank(), axis.title.y = element_blank()) 

## Create an appropriate dataframe with gross, imdb_score and genres for each movie
movies_and_genres <- cbind(gross = movies$gross, score = movies$imdb_score, movie_title = movies$movie_title, movies_and_genres, stringsAsFactors = FALSE)
## saving a full wide data frame to be used in later analysis
movies_and_genres_wide <- movies_and_genres
movies_and_genres <- melt(movies_and_genres, id = c("gross", "score", "movie_title"))
movies_and_genres$variable <- gsub("[.]", " ", movies_and_genres$variable)
movies_and_genres <- movies_and_genres[movies_and_genres$value == 1, ] 
movies_and_genres$value <- NULL
colnames(movies_and_genres) <- c("gross", "score", "movie_title", "genre")
movies_and_genres$genre <- factor(movies_and_genres$genre, levels = genres_sum$genre)
movies_and_genres <- movies_and_genres[complete.cases(movies_and_genres), ]
movies0 <- movies[movies$plot_keywords != "", ]
keywords <- c()
i <- 1
for (ins in movies0$plot_keywords){
    kw <- strsplit(ins, "[|]")
    if (length(kw) != 0){
        for (word in kw[[1]]){
            if (!(word %in% keywords)){
                keywords[i] <- word
                i = i + 1
            }
        }
    }
}
# Create a dataframe with logical values which 
# indiacte the keywords of each movie
movies0$plot_keywords <- strsplit(movies0$plot_keywords, "[|]")
keywords_idx <- movies0[, c("movie_title", "plot_keywords")]
i = 1
mat <- matrix(rep(0, (dim(movies0)[1] * length(keywords))), nrow = dim(movies0)[1])
for (word in keywords_idx$plot_keywords){
    idx <- which(keywords %in% word)
    mat[i, idx] <- 1
    i = i + 1
}
colnames(mat) <- keywords
movies_and_keywords <- data.frame(mat)

# Find how many movies belong in each keyword
sum <- rep(0, length(keywords))
for (i in 1:length(keywords)){
    sum[i] <- sum(movies_and_keywords[, i])
}
keywords_sum <- data.frame(keywords = factor(keywords), sum = sum)
keywords_sum <- keywords_sum[order(sum, decreasing = FALSE),]
keywords_sum$keywords <- factor(keywords_sum$keywords, levels = keywords_sum$keywords)
#keywords_sum <- keywords_sum[keywords_sum$sum > 39, ]
keywords_sum <- keywords_sum[(dim(keywords_sum)[1]-19):dim(keywords_sum)[1] ,]
movies_and_keywords <- cbind(gross = movies0$gross, score = movies0$imdb_score, movies_and_keywords, stringsAsFactors = FALSE)
movies_and_keywords <- melt(movies_and_keywords, id = c("gross", "score"))
movies_and_keywords$variable <- gsub("[.]", " ", movies_and_keywords$variable)
movies_and_keywords <- movies_and_keywords[movies_and_keywords$value == 1, ] 
movies_and_keywords$value <- NULL
colnames(movies_and_keywords) <- c("gross", "score", "keywords")
movies_and_keywords$keywords <- factor(movies_and_keywords$keywords, levels = keywords_sum$keywords)
movies_and_keywords <- movies_and_keywords[complete.cases(movies_and_keywords), ]
genres_model_gross <- lm(gross ~ genre, data = movies_and_genres)
#summary(genres_model_gross)
intercept1 <- lm(movies_and_genres$gross ~ genres_model_gross$fitted)$coefficients[[1]]

movies_and_genres_full <- data.frame(movies_and_genres[,c("gross","score","genre")], predicted_gross = genres_model_gross$fitted)
avg_gross <- movies_and_genres_full[, 3:4]
avg_gross <- avg_gross[!duplicated(avg_gross$genre), ]
avg_gross <- avg_gross[order(avg_gross$predicted_gross), ]
rownames(avg_gross) <- 1:dim(avg_gross)[1]
keywords_model_gross <- lm(gross ~ keywords, data = movies_and_keywords)
#summary(keywords_model_gross)
intercept3 <- lm(movies_and_keywords$gross ~ keywords_model_gross$fitted)$coefficients[[1]]

movies_and_keywords_full <- data.frame(movies_and_keywords, predicted_gross = keywords_model_gross$fitted)
avg_gross <- movies_and_keywords_full[, 3:4]
avg_gross <- avg_gross[!duplicated(avg_gross$keywords), ]
avg_gross <- avg_gross[order(avg_gross$predicted_gross), ]
rownames(avg_gross) <- 1:dim(avg_gross)[1]
```
For IMDB score prediction.

```{r, echo = FALSE, fig.width = 10, fig.align = "center"}
keywords_model_score <- lm(score ~ keywords, data = movies_and_keywords)
#summary(keywords_model_score)
intercept4 <- lm(movies_and_keywords$score ~ keywords_model_score$fitted)$coefficients[[1]]

movies_and_keywords_full <- data.frame(movies_and_keywords, predicted_score = keywords_model_score$fitted)
avg_score <- movies_and_keywords_full[, 3:4]
avg_score <- avg_score[!duplicated(avg_score$keywords), ]
avg_score <- avg_score[order(avg_score$predicted_score), ]
rownames(avg_score) <- 1:dim(avg_score)[1]

ggplot() + 
    geom_jitter(aes(x = movies_and_keywords_full$predicted_score, y = movies_and_keywords_full$score, 
                    colour = movies_and_keywords_full$keywords), alpha = 0.1) + 
    geom_abline(intercept = intercept4) + 
    geom_point(aes(x = avg_score$predicted_score, y = avg_score$predicted_score, colour = avg_score$keywords), 
               size = 2, shape = 21, stroke = 2) + 
    geom_text(aes(x = avg_score[seq(1, dim(avg_score)[1], 2), ]$predicted_score, 
                   y = avg_score[seq(1, dim(avg_score)[1], 2), ]$predicted_score, 
                   label = avg_score[seq(1, dim(avg_score)[1], 2), ]$keywords), 
               check_overlap = F, nudge_y = -0.2, nudge_x = 0, size = 3, angle = 45, fontface = "bold") + 
    geom_text(aes(x = avg_score[seq(2, dim(avg_score)[1], 2), ]$predicted_score, 
                   y = avg_score[seq(2, dim(avg_score)[1], 2), ]$predicted_score, 
                   label = avg_score[seq(2, dim(avg_score)[1], 2), ]$keywords), 
               check_overlap = F, nudge_y = 0.2, nudge_x = 0, size = 3, angle = 45, fontface = "bold") + 
    scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9)) +
    scale_x_continuous(breaks = c(6.1, 6.3, 6.5, 6.7)) +
    labs(title = "IMDB Score for different keywords", x = "Predicted IMDB Score", y = "Observed IMDB Score") +
    theme_bw() + 
    theme(legend.position = "None")
```

```{r, echo=FALSE}
# Boxplot of genres and imdb score
ggplot(movies_and_genres, aes(genre, score, fill = genre)) + 
    geom_boxplot() + stat_summary(fun.y=mean, colour="red", geom="point", size=1, show.legend = FALSE) +
    coord_flip() +
    labs(title = "IMDB score of movies by genre", x = "", y = "IMDB score") + 
    theme_bw() +
    theme(legend.position = "None") +
    theme(axis.ticks.y = element_blank(), panel.grid.major.y = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank()) +
    scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10), limits = c(1, 10))

# Boxplot of keywords and imdb score
ggplot(movies_and_keywords, aes(keywords, score, fill = keywords)) + 
    geom_boxplot() + stat_summary(fun.y=mean, colour="red", geom="point", size=1, show.legend = FALSE) +
    coord_flip() +
    labs(title = "IMDB score of movies by keyword", x = "", y = "") + 
    theme_bw() +
    theme(legend.position = "None") +
    theme(axis.ticks.y = element_blank(), panel.grid.major.y = element_blank(), axis.title.x = element_blank(), axis.title.x = element_blank()) +
    scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10), limits = c(1, 10))
```

For IMDB score prediction.

```{r, echo = FALSE, fig.width = 10, fig.align = "center"}
genres_model_score <- lm(score ~ genre, data = movies_and_genres)
#summary(genres_model_score)
intercept2 <- lm(movies_and_genres$score ~ genres_model_score$fitted)$coefficients[[1]]

movies_and_genres_full <- data.frame(movies_and_genres[,c("gross","score","genre")], predicted_score = genres_model_score$fitted)
avg_score <- movies_and_genres_full[, 3:4]
avg_score <- avg_score[!duplicated(avg_score$genre), ]
avg_score <- avg_score[order(avg_score$predicted_score), ]
rownames(avg_score) <- 1:dim(avg_score)[1]

ggplot() + 
    geom_jitter(aes(x = movies_and_genres_full$predicted_score, y = movies_and_genres_full$score, colour = movies_and_genres_full$genre), alpha = 0.1) + 
    geom_abline(intercept = intercept2) + 
    geom_point(aes(x = avg_score$predicted_score, y = avg_score$predicted_score, colour = avg_score$genre), 
               size = 2, shape = 21, stroke = 2) + 
    geom_text(aes(x = avg_score[seq(1, dim(avg_score)[1], 2), ]$predicted_score, 
                   y = avg_score[seq(1, dim(avg_score)[1], 2), ]$predicted_score, 
                   label = avg_score[seq(1, dim(avg_score)[1], 2), ]$genre), 
               check_overlap = F, nudge_y = -0.2, nudge_x = 0, size = 3, angle = 45, fontface = "bold") + 
    geom_text(aes(x = avg_score[seq(2, dim(avg_score)[1], 2), ]$predicted_score, 
                   y = avg_score[seq(2, dim(avg_score)[1], 2), ]$predicted_score, 
                   label = avg_score[seq(2, dim(avg_score)[1], 2), ]$genre), 
               check_overlap = F, nudge_y = 0.2, nudge_x = 0, size = 3, angle = 45, fontface = "bold") + 
    scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9)) +
    scale_x_continuous(breaks = c(6, 6.5, 7)) +
    labs(title = "IMDB Score for different genres", x = "Predicted IMDB Score", y = "Observed IMDB Score") +
    theme_bw() + 
    theme(legend.position = "None")
```



##Louise
```{r echo=FALSE}
##function to create human readable axis labels.
##code taken from publically available github to give human readable axis labels. (https://github.com/fdryan/R/blob/master/ggplot2_formatter.r)
human_numbers <- function(x = NULL, smbl =""){
  humanity <- function(y){             
    
    if (!is.na(y)){
      
       b <- round_any(abs(y) / 1e9, 0.1)
       m <- round_any(abs(y) / 1e6, 0.1)
       k <- round_any(abs(y) / 1e3, 0.1)
      
      if ( y >= 0 ){ 
        y_is_positive <- ""
      } else {
        y_is_positive <- "-"
      }
      
      if ( k < 1 ) {
        paste0(y_is_positive, smbl, y )
        } else if ( m < 1){
        paste0 (y_is_positive, smbl,  k , "k")
      } else if (b < 1){
        paste0 (y_is_positive, smbl, m ,"m")
      } else {
        paste0 (y_is_positive, smbl,  comma(b), "b")     
      }
    }
  }
  
  sapply(x,humanity)
}

human_num <- function(x){human_numbers(x, smbl = "")} 
human_usd <- function(x){human_numbers(x, smbl = "$")}
```

```{r}
directorssummary <- ddply(movies, ~ director_name,summarise,score_average=round(mean(imdb_score),2), gross_average=round(mean(gross),2), number_of_movies=length(director_name), director_facebook_likes = max(director_facebook_likes))

##sort by # of movies then imdb average score
sorteddirectorsummary <- arrange(directorssummary,desc(number_of_movies), desc(score_average))
##ensure that factors are in the order of number of movies, otherwise ggplot will default to alphabetical ordering in the graph
sorteddirectorsummary$director_name <- factor(sorteddirectorsummary$director_name, levels = sorteddirectorsummary$director_name[order(sorteddirectorsummary$number_of_movies)])
sorteddirectorsummary$more_than_one_movie <- rep.int(0, nrow(sorteddirectorsummary))
sorteddirectorsummary$more_than_one_movie[sorteddirectorsummary$number_of_movies > 1] <- 1
sorteddirectorsummary$more_than_ten_movies <- rep("Fewer than 10", nrow(sorteddirectorsummary))
sorteddirectorsummary$more_than_ten_movies[sorteddirectorsummary$number_of_movies>10] <- "10+"
movieswithdirectordata <- merge(movies, sorteddirectorsummary, by = "director_name")
movieswithdirectordata <- rename(movieswithdirectordata, c('number_of_movies'='dir_number_of_movies'))
```

```{r echo = FALSE, message = FALSE, fig.width = 3, fig.height=3, fig.align = "center" }
##plot for budget vs number of directors in sample
ggplot(movieswithdirectordata, aes(x = dir_number_of_movies, y = budget)) + 
    geom_jitter(alpha = 0.1, width = 1) +
    theme_bw() + 
    labs(x = "Number of movies by director in sample", y = "Budget") + 
    scale_y_continuous(label = human_usd) + 
    scale_x_continuous(breaks = c(0, 5, 10, 15, 20, 25, 30), limits = c(0, 30))
```

If we then split this out by decades, we can see the below:
```{r echo = FALSE}
##add the decades in as per Steven's code (note: potentially change into a function?)
movieswithdirectordata$decade <- sapply(movieswithdirectordata$title_year, FUN = convert_decade)
```

```{r echo = FALSE, fig.width = 10, fig.align = "center", warning = FALSE, message = FALSE}
ggplot(movieswithdirectordata, aes(x = dir_number_of_movies, y = gross, colour = decade)) + 
    geom_jitter(alpha = 0.1) + 
    scale_y_continuous(label = human_usd) + 
    scale_x_continuous(breaks = c(0, 5, 10, 15, 20, 25, 30), limits = c(0, 30)) +
    theme_bw() +
    labs(title = "Gross Revenue by\n number of movies per director\n by decades", x = "Number of movies\nper director", y = "Gross Revenue") + 
    facet_wrap(~ decade, nrow = 4) + 
    theme(legend.position = "None") + 
    geom_smooth(color="#000000")
```

As the number of data elements in the sample grows for each decade, a geomtrical pattern also seems to emerge, which is an upwards trend of gross revenue by movies per director up to about 10 movies per director, andthen a less clear trend for movies for movies with 10+ movies per director.

```{r echo=FALSE, message=FALSE, fig.width = 10, fig.align = "center"}
##plot for imdb rating vs # of movies
ggplot(movieswithdirectordata, aes(x = dir_number_of_movies, y = imdb_score)) + 
    geom_jitter(alpha = 0.2, width = 1) + 
    geom_smooth() +
    theme_bw() +
    labs(x = "Number of movies in sample", y = "IMDB Score") +
    scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10), limits = c(1, 10))
```

```{r echo=FALSE}
##cor test for imdb rating vs # of movies
pander(cor.test(movieswithdirectordata$imdb_score, movieswithdirectordata$dir_number_of_movies), caption="")
```

#To be removed- Notes and comments:

At the end of the assignment we should go back to:

* Check coding standards are consistent (and align with his recommendation - http://adv-r.had.co.nz/Style.html)
* Check language is consistent (tense / case)
* Convert ggplots into the same theme - colour scheme, fonts, etc.
* All assumptions have been noted in the appropriate sections
* We think we're roughly aiming for 20ish pages - tbd at a later stage
* Make sure we don't use language that implies causation when we can only infer correlation
* Double check each others' analyses for Simpson's paradox.
* Maybe put all libraries at the top?
