---
title: "Louise_Initial"
output: pdf_document
---
```{r setup}
library(ggplot2)
library(ggthemes)
library(plyr)
library(reshape2)
```


```{r loaddata}
movies <- read.csv("movie_metadata.csv", header = T, stringsAsFactors = F)
str(movies)
```

# Some initial plots

What countries make up the top 5000
```{r}
counted <- count(movies, 'country')
removednas <- counted[complete.cases(counted),]
sorted <- arrange(removednas,desc(freq))
head <- sorted[1:10,]
head$country <- factor(head$country, levels = head$country[order(head$freq)])

ggplot(head, aes(x=country, y=freq, fill=country, label=freq)) + geom_bar(stat="identity") + coord_flip() + geom_text(size=3,nudge_y=150) + guides(fill=FALSE) + scale_y_continuous(expand = c(0, 0)) +  expand_limits(y = c(0,1.08 * max(head$freq))) + ggtitle("Top 10 countries with multiple IMDB top 5000 movies") + labs(x = "", y="Frequency") 
```

What directors make up the top 5000
```{r}
movies$director_name[movies$director_name == ""] <- NA
counted <- count(movies, 'director_name')
removednas <- counted[complete.cases(counted),]
sorted <- arrange(removednas,desc(freq))
head <- sorted[1:20,]
head$director_name <- factor(head$director_name, levels = head$director_name[order(head$freq)])

ggplot(head, aes(x=director_name, y=freq, fill=director_name)) + geom_bar(stat="identity") + coord_flip() + guides(fill=FALSE) + scale_y_continuous(expand = c(0, 0)) + ggtitle("Top 20 directors with multiple IMDB top 5000 movies") + labs(x = "", y="Frequency") +  scale_y_continuous(expand = c(0,0)) +  expand_limits(y = c(0,1.05 * max(head$freq)))

dirpercentage <- round(100*length(sorted$freq[sorted$freq==1])/length(sorted$freq),2)

```

Approx  `r dirpercentage` of directors in the top 5000 have only 1 movie in the top 5000. So `r 100-dirpercentage` have more than 1.

##What is the rating frequency of the top 5000?
```{r}

##replace non top 10 countries with Other
movies$country_new <- movies$country
movies$country_new[!(movies$country %in% head$country)] <- "Other"

ggplot(movies, aes(x=imdb_score, color=country_new)) + geom_histogram(binwidth = 0.25)
```

Looks somewhat normal with a  bell curve.
Looks like there are more around whole and half numbers, than there are around x.25 / x.65s - this is likely due to the fact that users enter scores that are whole numbers only.

```{r}

## Have a look at NAs

  columnindex = integer()
  columns = character()
  nas = numeric()
  napercentage = numeric ()
  
for (k in 1:dim(movies)[2])
  {columnindex = c(columnindex,k)
  columns = c(columns,colnames(movies)[k])
  nas = c(nas,length(movies[[k]][is.na(movies[[k]])]))
  napercentage = c(napercentage,round(length(movies[[k]][is.na(movies[[k]])])/nrow(movies),4)*100)}
naanalysis <- data.frame(columnindex=columnindex,columns=columns,nas=nas,napercentage=napercentage)
naanalysis
```

## Do some cleaning

For fields with a low value of NAs, replace those values with the mean for that field.

```{r}
  #this doesn't work atm  
for (i in naanalysis$columnindex[naanalysis$napercentage > 0 & naanalysis$napercentage <1])
    {
    k = ncol(movies)+1
    movies[[k]] <- movies[[i]]
    movies[[k]][is.na(movies[[i]]==TRUE)] <- round(mean(movies[[i]], na.rm=TRUE),0)
    colnames(movies)[[k]] <- paste(colnames(movies[i]), "_clean", sep="")
}
```


## Make new correlation plot 
```{r corellationplot}
moviesnumeric<- movies[c("duration_clean","title_year","budget","imdb_score","aspect_ratio","movie_facebook_likes","actor_1_facebook_likes_clean","actor_2_facebook_likes_clean","actor_3_facebook_likes_clean","cast_total_facebook_likes","director_facebook_likes","num_user_for_reviews_clean","num_critic_for_reviews_clean","facenumber_in_poster_clean","gross")]
##used pairwise complete obs to handle NAs for fields with a high number of nas
corrmatrix <- round(cor(moviesnumeric, use="pairwise.complete.obs"),2)
meltedmovies<- melt(corrmatrix)
ggplot(data = meltedmovies, aes(x=Var1, y=Var2, fill=value)) + geom_tile() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(x="", y="") + ggtitle("Correlation Matrix")
```

There's a clear bundle of correlation for facebook likes (actors - cast which makes sense by design and cast - movies which makes sense as actors are likely to share their movies on facebook and encourage fans to like the movie, and fans of a movie are likely to allow like the actors in that movie.) Director facebook likes doesn't look like it's correlated with the rest. There are also quite a high percentage of director facebook like nas - this could suggest that directors aren't as active on facebook as actors.

There is another bundle of correlation for number of critics and users for reviews. These two variables also seem correlated with movie facebook likes. These could all be a proxy for public interest in the movie / number of people who saw the movie.

What is interesting is what "gross" and "imdb score" are correlated with as these are the "outcome" variables, that could be used to measure success.

Both of these look like they is correlated with movie facebook likes, and the number of user and critic reviews, with gross being more highly correlated than imdb_score. Again, these could all be a proxy for public interest in the movie / number of people who saw the movie. Imdb_score and gross also look like they may be correlated with each other.



###Some hypothesese it might be good to test

* Is gross correlated with imdb score (higher -> higher)
* Is gross - budget correlated with imdb score (higher -> higher)
* Is the sum of cast total facebook likes + movie facebook likes a predictor of gross / imdb score
* Is gross correlated with budget(higher -> higher)
* Is title year correlated with higher imdb score
* Is # of director imdb top 5000 movies correlated with gross / imdb score (need to think carefully about the maths and inference here)

##Some plots it might be nice to see
* Ratings broken down by genre (boxplot)
* Top actors in terms of facebook likes
* Ratings by year
* Budget to Gross
* Total number of facebook likes - histogram


##Notes and things to be removed


```{r eval = FALSE}
##replace NAs with mean values for analysis

movies$durationclean <- movies$duration
movies$durationclean[is.na(movies$duration)==TRUE] <- round(mean(movies$duration, na.rm=TRUE),0)

movies$facenumber_in_posterclean <- movies$facenumber_in_poster
movies$facenumber_in_posterclean[is.na(movies$facenumber_in_poster)==TRUE] <- round(mean(movies$facenumber_in_poster, na.rm=TRUE),0)

movies$num_critic_for_reviewsclean <- movies$num_critic_for_reviews
movies$num_critic_for_reviewsclean[is.na(movies$num_critic_for_reviews)==TRUE] <- round(mean(movies$num_critic_for_reviews, na.rm=TRUE),0)

```

